Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae

### 2020-11-23 introduction.tex
Some editing in the intro.

All multidimensional numerical simulations show the systematic appearance in time-frequency diagrams (or spectrograms) of a distinct and relatively narrow feature 
%
%{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} 
%
during the post-bounce evolution of the system, with frequency rising 
from about \unit[100]{Hz} up to a few kHz (at most) and a typical duration of \unit[0.5 -- 1]{s}. This feature has been interpreted as a continuously excited gravity mode (g-mode, see~\citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS~\citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015wra, Kuroda:2016, Andresen:2017}. 
In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
%promising {source} for 
ground-based GW interferometers.

 The {properties of} g-modes in hot {PNSs} have been studied since the 1990s
 %end of last century 
 {by means of linear perturbation analysis of background PNS models}. The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.}~\citep{McDermott:1983}. Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core~\citep{Reisenegger:1992}. Many subsequent works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation~\citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities~\citep{Dimmelmeier:2006}, phase transitions~\citep{Kruger:2015} and realistic equation of state~\citep{Camelio:2017}. {Only recently, there have been efforts to incorporate more suitable backgrounds based on numerical simulations in the computation of the mode structure and evolution~\citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.

Using results from 2D CCSN numerical simulations as a background~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included~\citep{Torres:2019a}).
%\tf{This last sentence might be improved.} \mab{MAB: What about: Computing the mode spectrum of the region within the shock (including the PNS and the post-shock region) consists in solving an eigenvalue problem using background simulations. Using results from 2D CCSN numerical simulations \citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}). } \pcd{[PCD: I am not convinced about the first sentence of mab's proposal. Maybe if we just delete "with respect to which to solve an eigenvalue problem" the sentence becomes more readable]}

@@

Many previous data analysis efforts focused on the reconstruction of the gravitational wave strain amplitude
without assuming a particular signal model~\citep{Summerscales:2008,Klimenko:2015ypf,CornLitt}. As numerical
simulations became more available other methods using catalogs of waveforms have been proposed to identify
the supernova explosion mechanism.
Among them, principal component analysis helps at reducing the complexity of CCSN waveform catalogs to fewer
parameters~\citep{Heng:2009,roever:09,Edwards:2014,powell:2016,powell:2017,Roma:2019kcd}.
The classification challenge is also well addressed with deep learning methods~\citep{astone:2018,Chan:2019fuz}.

>>>>

All multidimensional numerical simulations show the systematic appearance in time-frequency diagrams (or spectrograms) of a distinct and relatively narrow feature 
%
%{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} 
%
during the post-bounce evolution of the system, with frequency rising 
from about \unit[100]{Hz} up to a few kHz (at most) and a typical duration of \unit[0.5 -- 1]{s}. This feature has been interpreted as a continuously excited gravity mode (g-mode, see~\citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS~\citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015wra, Kuroda:2016, Andresen:2017}. 
In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for ground-based GW interferometers.

 The {properties of} g-modes in hot {PNSs} have been studied since the 1990s {by means of linear perturbation analysis of background PNS models}. The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.}~\citep{McDermott:1983}. Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core~\citep{Reisenegger:1992}. Many subsequent works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation~\citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities~\citep{Dimmelmeier:2006}, phase transitions~\citep{Kruger:2015} and realistic equation of state~\citep{Camelio:2017}. {Only recently, there have been efforts to incorporate more suitable backgrounds based on numerical simulations in the computation of the mode structure and evolution~\citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.

Using results from 2D CCSN numerical simulations as a background~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included~\citep{Torres:2019a}).

@@

Previous data analysis efforts have focused on the reconstruction of the GW strain amplitude
without assuming a particular signal model~\citep{Summerscales:2008,Klimenko:2015ypf,CornLitt}. As the amount of numerical
simulations increased other methods using waveform catalogs have been proposed to identify
the supernova explosion mechanism.
Among them, principal component analysis helps at reducing the complexity of CCSN waveforms to fewer
parameters~\citep{Heng:2009,roever:09,Edwards:2014,powell:2016,powell:2017,Roma:2019kcd}.
The classification challenge is also well addressed with deep learning methods~\citep{astone:2018,Chan:2019fuz}.
